---
title: "DS_ML_HousePrice"
author: "Tahani AlHarbi"
date: "12/3/2020"
output: html_document
---

data from: https://www.kaggle.com/shree1992/housedata
the data contains information  about houses in USA 
it contains 18 variables and 4600 observation
my main goal is to building a model to predict the price of a house
based on specific features of the house.



```{r setup, include=FALSE}
knitr::opts_chunk$set(include = FALSE)
```

```{r}
# Helper packages
library(dplyr)     # for data manipulation
library(ggplot2)   # for awesome graphics

# Modeling process packages
library(rsample)   # for resampling procedures
library(caret)     # for resampling and model training
library(h2o)       # for resampling and model training

# h2o set-up 
h2o.no_progress()  # turn off h2o progress bars
h2o.init()         # launch h2o
```








```{r}
house <- read.csv("data/data.csv")

```



```{r}
house$statezip <- as.factor(house$statezip)
house$city <- as.factor(house$city)
house$street <- as.factor(house$street)
```



```{r}
#change bathrooms, bedrooms into numeric 

house$price <- as.numeric(house$price)
house$bedrooms <- as.numeric(house$bedrooms)
house$bathrooms <- as.numeric(house$bathrooms)

```


```{r}

#we have statezip so we do not need for columns city, street 
#drop column of country because it has just one value 
#drop column of date beacuse it isn't necessary 


house <- subset( house , select = -c( date,country, street,city) )

str(house)


```


#### Data:
 

```{r}
head(house)
```




```{r}
tail(house)
```



```{r}
# initial dimension
dim(house)
# response variable
head(house$price)
```


#### **set h2o environment:**

```{r}
house.h2o <- as.h2o(house)
```

#### **Data Splitting**:

```{r}
# Stratified sampling with the rsample package
set.seed(123) 
split <- initial_split(house, prop = 0.7, 
                       strata = "price")
house_train  <- training(split)
house_test   <- testing(split)
```

#### **To make sure training and test set have the same distributions:**

```{r}
# Do the distributions line up? 
ggplot(house_train, aes(x = price)) + 
  geom_line(stat = "density", 
            trim = TRUE) + 
  geom_line(data = house_test, 
            stat = "density", 
            trim = TRUE, col = "red")
```


#### **Missing valuse:** 

visualize missing values 



```{r}
house %>%
  is.na() %>%
  reshape2::melt() %>%
  ggplot(aes(Var2, Var1, fill=value)) + 
    geom_raster() + 
    coord_flip() +
    scale_y_continuous(NULL, expand = c(0, 0)) +
    scale_fill_grey(name = "", 
                    labels = c("Present", 
                               "Missing")) +
    xlab("Observation") +
    theme(axis.text.y  = element_text(size = 4))
```


obviously we don not have any missing values. 



```{r}
# Specify resampling strategy
cv <- trainControl(
  method = "repeatedcv", 
  number = 10, 
  repeats = 5
)

# Create grid of hyperparameter values
hyper_grid <- expand.grid(k = seq(2, 25, by = 1))

# Tune a knn model using grid search
knn_fit <- train(
  price ~ ., 
  data = house_train, 
  method = "knn", 
  trControl = cv, 
  tuneGrid = hyper_grid,
  metric = "RMSE"
)
```




```{r}
# Print and plot the CV results
knn_fit
```



```{r}
ggplot(knn_fit)
```

